---
title: "Education_monitoring_preprocess"
author: "Henryzyguo"
date: "11/28/2017"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(plyr)
library(dplyr)
library(tidyr)
library(reshape2)
```


```{r read data, include=FALSE, warning= FALSE}
enr_age <- read.csv('enr_by_agegrd.csv')
staff_ed <- read.csv('staff_by_ed.csv')
ent_new <- read.csv('new_entrants.csv')
# schage_pop <- read.csv('schage_population.csv')
# staff_qf <- read.csv('staff_by_qlf.csv')
stu_grd9 <- read.csv('Grd_0910Students.csv')
acad_grd9 <- read.csv('Grd_0910Academics.csv')
acad_12 <- read.csv("Grd1112_Academics.csv")
acad_overview <- read.csv('aca_overview.csv')
```

##Students Stats Plots
suppose the audience has selected school 'Addibre', the graph would look like below:
```{r plot1, warning= FALSE}
enr_age<-filter(enr_age,SCHOOL=="Addibre")
G1<-select(enr_age,starts_with("G1"),Acad_yr)
#G1F<-select(enr_age,starts_with("G1F"),Acad_yr)
G1<-arrange(G1,desc(Acad_yr))
G1$Acad_yr <-NULL 

rownames(G1)<-(2004:2010)
G1<-as.table(t(G1))

dat2 <- filter(melt(G1),value!=0)
sex1<-substr(dat2$Var1,3,3)
age1<-substr(dat2$Var1,4,5)
dat2<-cbind(dat2,sex1,age1)
dat2$age1 <- as.numeric(as.character(dat2$age1))
ggplot(dat2, aes(x=sex1, y=value, fill=factor(age1))) + geom_bar(stat="identity") +labs(x="year",y="person number",fill="age",title="Age component each year")+
scale_fill_manual(values=c("indianred", "orange", "gold","skyblue4","light blue","blue"))+theme(plot.title = element_text(hjust = 0.5))+facet_grid(.~Var2)
```

Graph for new entrants' distribution by gender and age, still by school and the default is "Addibre" 
```{r pressure, echo=FALSE, warning= FALSE}
#filter
# ent_new <- read.csv('new_entrants.csv')
ent_new <-filter(ent_new,SCHOOL=="Addibre")
ent_new <-select(ent_new,4:12)
rownames(ent_new)<-c(2006,2010)

ent_new <-select(ent_new,-Acad_yr)
#melt
dat3 <- filter(melt(t(ent_new)),value!=0)

ggplot(dat3, aes(x=Var2, y=value, fill=as.factor(Var1))) + geom_bar(stat="identity",width = 0.5) +labs(x="year",y="person number",fill="age",title="new entry Age conponent each year")+
scale_fill_manual(values=c("indianred", "orange", "gold","skyblue4","light blue","blue"))+theme(plot.title = element_text(hjust = 0.5))

ent_new <- read.csv('new_entrants.csv')
ent_new <-filter(ent_new,SCHOOL=="Addibre")
ent_new <-select(ent_new,4:12)
rownames(ent_new)<-c(2006,2010)
ent_new <-select(ent_new,-Acad_yr)
#melt
dat3 <- melt(t(ent_new))
dat3 %>% separate(Var1,c("male","age"))
sex<-substr(dat3$Var1,1,1)
age<-substr(dat3$Var1,2,4)
dat3<-cbind(dat3,sex,age)
ggplot(dat3, aes(x=age, y=value,fill=as.factor(sex))) + geom_bar(stat="identity") +labs(x="age",y="person number",title="New entrant each year",fill="sex")+
scale_fill_manual(values=c("indianred", "orange", "gold","skyblue4","light blue","blue"))+theme(plot.title = element_text(hjust = 0.5))+facet_grid(.~Var2)
```


Save staffed.expanded to RData
```{r process_teachers_data, echo=FALSE}
# staff_ed <- read.csv('staff_by_ed.csv')
# replace all null values in df
staff_ed[is.na(staff_ed)] <- 0
# new column is sum of male and female
for(i in 3:24){
  staff_ed[,(47+i)] <- staff_ed[,(i*2-1)]+staff_ed[,(i*2)]
  colnames(staff_ed)[47+i]<-gsub('M','',colnames(staff_ed)[i*2-1])
}
# delete all columns of sex, directory
staff_ed <- staff_ed[,-c(5:48)]
staff_ed <- subset(staff_ed,select = -Qlf_Code)
# expand by grades 1-4 or 5-8
staff_ed14 <- staff_ed[,c(1:5,7,9,11,13,15,17,19,21,23,25)]
staff_ed58 <- staff_ed[,c(1:4,6,8,10,12,14,16,18,20,22,24,26)]
staff_ed14$grade <- 'Grade 1 to 4'
staff_ed58$grade <- 'Grade 5 to 8'
colnames(staff_ed14)[5:14]<-gsub('_1_4','',colnames(staff_ed14)[5:14])
colnames(staff_ed14)[15] <- 'NR'
colnames(staff_ed58)[5:15]<-colnames(staff_ed14)[5:15]
staff_ednew <- rbind(staff_ed14,staff_ed58)
remove(staff_ed14,staff_ed58)
# create columns for saving Qualification levels, not recording the teachers whose qualifications are not recorded
for(i in 5:14){
  staff_ednew[i+12] <- as.numeric(gsub('Q','',colnames(staff_ednew)[i]))
}
# expand the first column Q1
staffed.expanded <- staff_ednew[rep(row.names(staff_ednew), staff_ednew[,5]), c(1:4,17,16)]
colnames(staffed.expanded)[5] <- 'Qualification'
# expand all qf columns
for(i in 6:14){
  staffed.new <- staff_ednew[rep(row.names(staff_ednew), staff_ednew[,i]), c(1:4,i+12,16)]
  colnames(staffed.new)[5] <- 'Qualification'
  staffed.expanded <- rbind(staffed.expanded,staffed.new)
}
staffed.expanded$Qualification[staffed.expanded$Qualification==14]<-11
remove(staff_ednew,staffed.new)
```

##Teacher Certification Visualization
A violin plot to present teachers' certification level
can also be controlled with a 'school' button
```{r plot teacher}
# teacher
# caculate mean
staffed.expanded <- staffed.expanded[(staffed.expanded$SCHOOL=="Welel"),]
staff_means <- ddply(staffed.expanded, "grade", summarise, mean_Q = mean(Qualification))
violin <- ggplot(staffed.expanded, aes(x=grade, y=Qualification, fill=grade, color=grade)) + 
  geom_violin(show.legend = TRUE) +
  labs(title="Teachers' Qualification by Grades",x="Teaching Grades", y = "Years of Education after Grade 10") +
  facet_grid(~grade,scale ="free")
violin + 
  # stat_summary(fun.y = 'median', geom="point",color='blue',size = 2,show.legend = FALSE) +
  geom_hline(data=staff_means, aes(yintercept=mean_Q), color="orange", linetype="dashed",show.legend = TRUE)+
  scale_fill_manual(values=c("indianred", "gold"))+
  scale_color_manual(values=c("indianred", "gold"))
  cat("The mean after-primary-education years of Grade 1-4 teachers is",staff_means[1,2],"years\n")
  cat("The mean after-primary-education years of Grade 5-8 teachers is",staff_means[2,2],"years")
  
  
  staffed.expanded2 <- staffed.expanded[(staffed.expanded$SCHOOL=="Addibre"),]
  staff_means <- mean(staffed.expanded2$Qualification, na.rm = TRUE)
  ggplot(data = staffed.expanded2, aes(staffed.expanded2$Qualification, fill=grade, color=grade)) + 
        geom_histogram(show.legend = TRUE) +
        labs(title="Teachers' Qualification by Grades",x="Teaching Grades", y = "Years of Education after Grade 10") +
        scale_fill_manual(values=c("indianred", "gold"))+
        scale_color_manual(values=c("indianred", "gold"))
```

Student/Teacher Ratio.
The problem is that we only know 3 schools' total number of students while one of them has no teacher information.
So the visualization only includes 2 schools.
```{r plot_student, warning=FALSE}
enr_age <- read.csv('enr_by_agegrd.csv')
enr_age$Std_Boy <- rowSums(enr_age[grep("M",colnames(enr_age))])
enr_age$Std_Girl <- rowSums(enr_age[grep("F",colnames(enr_age))])
# add a total number of teachers
staff_ed$total <- rowSums(staff_ed[,5:26])
tch_std <- enr_age[,c(3,4,183,184,185)] %>% inner_join(staff_ed[,c(3,4,27)], by = c("SCHOOL","Acad_yr"))
# tch_std <- melt(tch_std, id=c("SCHOOL","Acad_yr","Tch_Sum"))
p<-ggplot(data=tch_std, aes(x=SCHOOL, y=Tab_16/total, fill=SCHOOL)) +
  geom_bar(stat="identity",width=0.4)+
  labs(title="Students/Teachers Ratio in 2006",x="School Name", y = "Average Student Number of A Teacher") 
# Horizontal bar plot
p + coord_flip()
```


##Academic Visualization
grd9 is the dataset for academics of Grade9.
The grpah provided is based on the hypothesis that a user selected year 10 and exam Tigrigna.Sem1.G9
Overall passing rate is calculated and will be presented on the final website
```{r acad_distribution, message=FALSE, warning=FALSE}
# find the correct column
# acad_12[grepl('11',colnames(acad_12)),]

stu_grd9 <- read.csv('Grd_0910Students.csv')
acad_grd9 <- read.csv('Grd_0910Academics.csv')
colnames(stu_grd9)[13]<-'Stud.ID'
stu_grd9 <- stu_grd9[,colSums(is.na(stu_grd9))<nrow(stu_grd9)]
acad_grd9 <- acad_grd9[,colSums(is.na(acad_grd9))<nrow(acad_grd9)]
grd9 <- acad_grd9 %>% inner_join(stu_grd9[,c(6,7,8,9,11)], by = "Stud.ID")
grd9$Sex2[grd9$Sex=="M"] <- "Male"
grd9$Sex2[grd9$Sex=="F"] <- "Female"
grd9$Sex <- grd9$Sex2
# grd9$Age_startGrd9 <- as.factor(grd9$Age-grd9$Grade+9)
# choose rows grades are not NA and columns
selected <- grd9 %>% subset(Tigrigna.Sem1.G9 !='NA'&Acad.yr.G9=='9/14/10 00:00',select = c(Acad.yr.G9,Tigrigna.Sem1.G9,Sex))
# create a table to save pass rate
pass <- ddply(selected, "Sex", summarise, passrate = 
                length(selected$Tigrigna.Sem1.G9[selected$Tigrigna.Sem1.G9>=50])/length(selected$Tigrigna.Sem1.G9))
print(cat('passing rate is',pass[1,2]))
p  <- ggplot(selected, aes(Tigrigna.Sem1.G9, fill=Sex, color=Sex))+
  scale_color_manual(values=c("indianred", "gold"))
p + geom_density(alpha=0.5)+
  scale_fill_manual(values=c("indianred", "gold"))+
  geom_vline(data=selected, aes(xintercept=50), color="red", linetype="dashed",show.legend = TRUE)+
  labs(title=paste("Students Grade Distribution,",'passing rate is',unlist(pass[1,2]),sep=" "),x="Grades", y = "Student Percentage Density") 





  acad_overview <- read.csv('aca_overview.csv')
    acad_overview <- filter(acad_overview,Semester=="1")
    acad_overview <- select(acad_overview,-Semester)
    #rownames(acad_overview) <-c(9:12)
    acad_overview <- melt(acad_overview)
    #rownames(acad_overview)<-acad_overview$grade
    #acad_overview<-t(acad_overview)
    ggplot(acad_overview, aes(x=variable,y=value,fill=variable)) + geom_bar(stat="identity")+
      scale_fill_manual(values=c("indianred", "orange", "gold","skyblue4","light blue","blue"))+
      labs(x="grade",y="score",title="Academic overview each year",fill="subject")+
      theme(plot.title = element_text(hjust = 0.5))+
      facet_grid(.~reorder(grade,value))
```


```{r save_tables}
ent_new[is.na(ent_new)] <- 0
staff_ed[is.na(staff_ed)] <- 0
saveRDS(ent_new,"ent_new.RData")
saveRDS(staff_ed,"staff_ed.RData")
saveRDS(enr_age,"enr_age.RData")

grd12 <- acad_12[,colSums(is.na(acad_12))<nrow(acad_12)]
colnames(grd12) <- gsub("Tig","Tigrigna",colnames(grd12))
colnames(grd12) <- gsub("Amh","Amharic",colnames(grd12))
colnames(grd12) <- gsub("Eng","English",colnames(grd12))
colnames(grd12) <- gsub("Chem","Chemistry",colnames(grd12))
colnames(grd12) <- gsub("Bio","Biology",colnames(grd12))
colnames(grd12) <- gsub("Phy","Physics",colnames(grd12))
colnames(grd12) <- gsub("Geo","Geography",colnames(grd12))
colnames(grd12) <- gsub("His","History",colnames(grd12))
colnames(grd12) <- gsub("Eco","Economics",colnames(grd12))
colnames(grd12) <- gsub("Buss","Bussine",colnames(grd12))

# colnames(grd12)[grepl('11',colnames(acad_12))]
# acad_12[grepl('11',colnames(acad_12)),]

grd9 <- grd9[,]
saveRDS(grd9[,-c(5:7,37,39)],"grd9.RData")
saveRDS(grd12[,-c(5,39,40)],"grd12.RData")

saveRDS(staffed.expanded,"staffed.expanded.RData")
saveRDS(acad_overview,"aca_overview.RData")
saveRDS(staff_ednew,"staff_ednew.RData")

enr_age <- readRDS("enr_age.Rdata")
staff_ed <- readRDS("staff_ed.Rdata")
ent_new <-readRDS("ent_new.RData")
grd9 <- readRDS("grd9.RData")
grd12 <- readRDS("grd12.RData")
staffed.expanded <- readRDS("staffed.expanded.RData")
staff_ednew <- readRDS("staff_ednew.RData")
acad_overview <- readRDS("aca_overview.RData")


```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
